\section{Code Generation}
This section describes the methodology and implementation of the code generation aspect of the compiler for HCL. 

During generation the HCL code is translated into C++, which is subsequently compiled using a C++ compiler.
If an Arduino board is connected to the computer, the compiler generated code that is specific to the Arduino language. 

The overall code generation is done within the \texttt{ProgramGenerator} class.
The class' responsibility is limited to calling the three predominant generators, namely \texttt{CodeGenerator}, \texttt{TypeGenerator} and \texttt{MainGenerator}.
Other than calling the other classes, the \texttt{ProgramGenerator} class also adds return code to the AST to make sure that the program returns when terminated. 

Snippet \ref{lis:programGen} shows the \textit{generate} function from \texttt{ProgramGenerator}.

\begin{lstlisting}[language=Kotlin,label=lis:programGen,caption=The implementation of \textit{generate} in \texttt{ProgramGenerator}.]
override fun generate(ast: AbstractSyntaxTree): List<FilePair> = 
	listOf(
		HelperHeaders.constList,
		HelperHeaders.ftoa,
		FilePair("builtin.h", CodeGenerator().generate(ast.builtins())),
		FilePair("types.h", TypeGenerator().generate(ast)),
		FilePair("main.cpp", MainGenerator().generate(ast.addReturnCode().notBuiltins()))
)
\end{lstlisting}

The function first creates header files for the help-classes \textit{constList}(3) and \textit{ftoa}(4), which allows for the use of the list-type in C++, and for converting doubles to text. 
The function then creates the header file for the built-in functions of the language(5).
When the \textit{generate} function within the \texttt{CodeGenerator} class is called with the built-in functions from the language, it formats the functions to resemble valid C++ function declarations, whilst also adding descriptive comments of its functionality.
The resulting text is written to a file that serves as a header-file for the final program. \\
The next line of code (6) calls the \textit{generate} function from the \texttt{TypeGenerator} class.
This class is responsible for formatting tuples in the language by creating equivalent \textit{struct} declarations in C++.
The functions fetches all tuples within the inputted AST, formats them, and writes the result of the formatting to a header file.
Before returning, the function also generates functions for each tuple found.
These generated functions allows the tuples to be printed using the \textit{toText} function, and will also allow the tuples to return a specific element from within themselves.\\
The last function called is the \texttt{MainGenerator} class' \textit{generate} function. 
This last \textit{generate} function is responsible for wrapping and setting up Arduino's \texttt{main}, \texttt{setup} and \texttt{loop} functions.
Before generation a return code is added to the AST, to ensure that the program has a final return.
Built-in functions are not added to generation, as these has already been handled. 

The \texttt{MainGenerator}'s \textit{generate} function is described in detail in snippet \ref{lis:mainGen}.
\begin{lstlisting}[language=Kotlin,label=lis:mainGen,caption=The implementation of \textit{generate} in \texttt{MainGenerator}.]
override fun generate(ast: AbstractSyntaxTree): String {
	val stringBuilder = StringBuilder(mainHeader)
	val declarations = ast.genFromFilterWithMap ({ it.isDecl }, 
	{
		// Find all declarations	
	})
	val setup = ast.genFromFilterWithMap ({ !it.isLoop && !it.isDecl }, 
	{
	    // Assign all declarations
	})
	val loop = ast.genForLoop()
	stringBuilder.appendln(declarations)
	stringBuilder.appendln(setup.wrapSetup())
	stringBuilder.appendln(loop.wrapLoop())

	stringBuilder.appendln(("setup();\n" +
	(if (loop.isNotBlank()) "while(1) { loop(); }\n" else "")).wrapMain()
	)
		return stringBuilder.toString()
	}
\end{lstlisting}
(2) Add all header files and namespaces.\\ 
(3) Find all declarations in source code.\\
(7) Find all commands in source code that is not declarations or associated with the \textbf{loop} function.\\
(10) Find all commands that are associated with the \textbf{loop} function.\\
(12) Add all declarations to output file.\\
(13) Wrap and add \textbf{setup} function to file.\\
(14) Wrap and add \textbf{loop} function to file.\\
(16) Wrap \textbf{main} function and add call to \textbf{setup} function.\\
if \textbf{loop} function is not empty, a call to it is also added to the \textbf{main} function.\\
(19) Source code is generated and returned from function. 

