% !TeX root = ../main.tex
\section{Parser implementation}
\label{sec:parser_impl}
\label{parserImplemention}
In the following section, the parser implementation details is touched upon.

Due to the nature of HCL, the parser can be implemented as a single pass parser. 
This is because the user may not access any variables that has not yet been declared. 
As HCL is a LL(K) language, an LL(k) recursive decent parser is a reasonable choice.

Both syntactical and contextual analysis is done within the parser. 
This means that there is no additional post parser type-checking or contextual analysis step. 
A major reason for doing in line contextual analysis is the way functions are invoked in HCL. 
The only way to know if an identifier is a function call, is by analyzing the types. 
This means that to create a proper AST, the type-checking must be done in line. 
It could also potentially be done afterwards, but that would require quite a large amount of post processing, which would be a lot more complex, than just doing type-checking within the parser. 
This also means that the outputted AST is in fact already a decorated syntax tree. 
For instance, function calls already has information on its return type, even though the function may have multiple return types based on the overloading of the function.

\subsection{Implementing a recursive decent parser}
A LL(K) recursive decent parser will do lookaheads until it is aware what branch should be parsed.
This can be figured out from the first and follow sets, that can be produced from the CFG.
Once it determines which non-terminal to parse it will call a method that attempts to parse that part.
The method can then accept terminals or parse further non-terminals.
It should be the responsibility of the accept method to advance the lexical token stream.

Below is a quick and concrete example, where the parser generates a token stream of the following HCL source code:
\begin{lstlisting}[language=HCL,label=lis:typedclsParserNum,firstnumber=1]
num x = 5
\end{lstlisting}

\begin{enumerate}
	\item Type.Number
	\item Identifier("x")
	\item SpecialChar.Equals
	\item Literal.Number(5.0)
	\item SpecialChar.EndOfLine
\end{enumerate}

The parser will see that the first token is a number type. 
Since only declarations have types in its first set, the parser now knows that it should parse a declaration.

The EBNF reveals that a declaration is:
\begin{itemize}
	\item Non-terminal Type
	\item Terminal Identifier
	\item Optional
	\begin{itemize}
		\item Terminal Equals
		\item Non-terminal Expression
	\end{itemize}
\end{itemize}
\textbf{Deciding what to parse}\\
The sealed classes, along with the pattern matching of Kotlin, are excellent tools for deciding what should be parsed. 
In snippet \ref{lis:parseCommand} a subset of the parse command method is shown. 

\begin{lstlisting}[language=Kotlin,label=lis:parseCommand,caption=A simplified version of the parse declaration method from the parser.]
private fun parseCommand(): AstNode.Command {
    val command = when (current.token) {
        is Token.Type -> parseDeclaration()
        is Token.Identifier ->
            if (peek().token == Token.SpecialChar.Equals) {
                parseAssignment()
            } else parseExpression()
        Token.Return -> parseReturnStatement()
        else parseExpression()

...

    flushNewLine()
    return command
}
\end{lstlisting}
Snippet \ref{lis:parseCommand} uses the functional functionality of Kotlin, to set the val\footnote{val is a read only variable} \textit{\textbf{command}} to a specific value based on the type of the current token (\textit{\textbf{current.token}}). 
This method calls other methods based on the type and based on specific rules in the HCL language. 
For instance all statements starting with a "type" is a declaration, as seen in the CFG, therefore the \textit{\textbf{parseDeclaration}} method is called, and the returned value of this method is then going to be assigned to the command value.
	
To simplify the parser and increase readability, certain helper methods have been defined. 
For instance the \textit{\textbf{accept<T>}} method. 
\begin{lstlisting}[language=Kotlin,label={lis:acceptMethod},caption=The definition of the accept method on the parser]
private inline fun<reified T: Token> accept(): T {
	val token = current.token
	moveNext()
	if (token is T) {
		return token
	} else {
		wrongTokenTypeError(T::class.simpleName!!, token)
	}
}
\end{lstlisting}
Snippet \ref{lis:acceptMethod} shows how the accept method is defined.
This method is used to pop a token from the token stream, which is done by the \textit{\textbf{moveNext()}} method, if the token has the expected type.
This is used for getting the current token if it's of the correct type. 
This can also be used even if the data of the token is irrelevant, for instance shown in snippet \ref{lis:acceptMethodHCLExample}, the accept method would be used to pop the equals-symbol. 
However it's important that the equals symbol is present, as the code would not make sense without it. 
The accept method would then trow an error if another token would be present.

\begin{lstlisting}[language=HCL,label={lis:acceptMethodHCLExample},caption={An example of a declaration in HCL, where the accept method is used by the parser.}]
num a = 5
\end{lstlisting}
Snippet \ref{lis:acceptMethodHCLExample} is a declaration, and these are handled by the \textit{\textbf{parseDeclaration}}, which is implemented as seen in snippet \ref{lis:parseDeclaration}.

This is a great example of how parsing is done on a specific type of statement.
Do note, the final parse declaration method is a little more complex due to type inference and type checking.

\begin{lstlisting}[language=Kotlin,label=lis:parseDeclaration,caption=A simplified version of the parse declaration method from the parser.]
private fun parseDeclaration(): AstNode.Command.Declaration {
    val type = parseType()
    val identifierToken = accept<Token.Identifier>()
    val identifier = AstNode.Command.Expression.Value.Identifier(identifierToken.value)
    val expression = if (current.token == Token.SpecialChar.Equals) {
	    accept<Token.SpecialChar.Equals>()
	    parseExpression()
    } else null
    return AstNode.Command.Declaration(type, identifier, expression)
}
\end{lstlisting}

The \textbf{\textit{parseDeclaration}} method, as shown in snippet \ref{lis:parseDeclaration}, will first save the type in the \textbf{\textit{type}} val, and then accept an identifier and save it to an instance of Identifier, a subclass of the AstNode class. 
The method then checks whether the new current symbol is an equals-symbol - if that's the case the declaration is followed by an expression - which means it's an declaration with an assignment otherwise the expression is set to null. 
The method then returns a Declaration, which is a subclass of AstNode.  

It is slightly more complex to parse an expression, since an expression in HCL may be either a function call or any kind of value.
This means that there are a lot of options for the parser when parsing expressions.
The parser first tries to parse a potential function call using the \textbf{\textit{parsePotentialFunctionCall}} method.

\begin{lstlisting}[language=Kotlin,label=lis:parsePotCall,caption=A simplified version of the parsePotentialFunctionCall method from the parser.]
private fun parsePotentialFunctionCall(expression: AstExpression? = null): AstExpression =
    when (current.token) {
        is Token.SpecialChar.BlockStart -> {
            val upcomingId = getUpcomingIdentifierNameForLambda()
            if (upcomingId != null) {
                val symbol = retrieveSymbol(upcomingId)
                val lambdaParameter = getLambdaParameter(symbol.functions, 0)
                parsePotentialFunctionCall(lambdaParameter)
            } else {
                val lambda = parseLambdaBody(emptyList())
                AstNode.Command.Expression.LambdaExpression(listOf(), lambda.type, lambda.lambdaBody)
            }
        is Token.Identifier -> {
            val token = accept<Token.Identifier>()
            val symbol = retrieveSymbol(token.value)
            if (symbol.isFunctions) {
        	    ...parse needed arguments...
        	    AstNode.Command.Expression.FunctionCall(
        	    AstNode.Command.Expression.Value.Identifier(token.value),
        	    listOf(expression) + secondaryArguments)
            } else expression
        } else -> expression
\end{lstlisting}

As seen in snippet \ref{lis:parsePotCall}, the method first determines if the current token is a lambda expression, signified by a \textit{blockStart-token} (3).
Should it be a lambda expression, the parser determines whether or not there is a functions call within the lambda expression that needs to be parsed as well (7).
If there is not a function call within the lambda expression, the lambda expression itself is returned from the method (10). 
Should the token be an identifier, the parser determines if the identifier is associated with a function, in which case it is determined that a function call is being made (16).
The method looks for the needed arguments for the function, and parse them as expressions.
The method then returns the function call with its corresponding arguments (18).
If the identifier does not signify a function, it is then parsed as an expression that is not a function call.

All expressions that are determined to not be function calls, are parsed by the method \textbf{\textit{parseExpressionAtomic}} seen in snippet \ref{lis:parseExpr}.

\begin{lstlisting}[language=Kotlin,label=lis:parseExpr,caption=A simplified version of the parseExpressionAtomic method from the parser.]
private fun parseExpressionAtomic(): AstExpression =
    when (current.token) {
        Token.SpecialChar.SquareBracketStart -> parseListDeclaration()
        Token.SpecialChar.ParenthesesStart -> {
            if (isLambdaParameters()) {
                parseLambdaDeclaration()
            } else {
                if(upcomingTuple()){
                    parseTupleExpression()
                } else {
                    parseExpression()
                }
            }
        }
        Token.Literal -> acceptLiteral()
        Token.SpecialChar.Colon -> {
        	accept<Token.SpecialChar.Colon>()
        	val token = accept<Token.Identifier>()
        	retrieveSymbol(token.value).handle(/*handle token as function*/)
        }
        Token.Identifier -> {
            val token = accept<Token.Identifier>()
            retrieveSymbol(token.value).handle(/*Handle identifier*/)
        }
        else -> unexpectedTokenError(current.token)
    }
            
\end{lstlisting}

The method itself determines the type of the expression that is to be parsed, and then parses it atomically.
All expressions that begin with the '\texttt{[}' character are determined to be list declarations and are parsed as such (3).\\
Expressions beginning with a parenthesis can either be a lambda declaration (6), a tuple (9) or simply an expression encapsulated in parentheses (11).\\
Literals are simply accepted (15).\\
Expressions beginning with a colon can only be functions passed as arguments, and are handled accordingly (16).\\
Identifiers can either signify variables or function calls.
However, as potential function calls have already been handled at this point, the identifier is determined to signify a variable and can be parsed accordingly (21).
If no eligible tokens has been found at this point in parsing, it is determined that a syntax error has occurred and the parser throws an exception (25).

% we may also need to document some of the complex parts of the parser, but we will wait with that, until it is actually more or less completed.
\input{4.Solution/SymbolTable.tex}
